Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
basic bandwidth

\section{Experimental Validation}
\label{sec:expe}
In this section, we describe a series of numerical experiments to confirm 
the performance predictions of Section~\ref{sec:model}. To better
understand our results, we first perform some bandwidth measurements, followed
by the actual SPmV.  \todo{We might be criticized for not comparing with other work.}

@@

\def\ww{.13\textwidth}
\begin{figure*}[tbh]
  \begin{center}
    \subfigure[Supercompact.]{%
      \label{fig:supercompact}
      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
    \subfigure[Compact.]{%
      \label{fig:compact}
      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
    \subfigure[Random.]{%
      \label{fig:random}
      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
    \subfigure[2D, no RCM.]{%
      \label{fig:rbf2dnorcm}
      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
    \subfigure[2D, RCM.]{%
      \label{fig:rbf2drcm}
      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
    \subfigure[3D, RCM.]{%
      \label{fig:rbf3dnorcm} 
      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
    \subfigure[3D, RCM.]{%
      \label{fig:rbf3drcm}
      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
  \end{center}

@@

The other two types of matrices are used in derivative stencils in 2D
and 3D RBF-FD calculations (Figure~\ref{fig:rbf2dnorcm} pictures a
$32$-points stencil of a $23^2$ 2D grid and
Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
grid). We also apply a Cuthill-Mckee reordering to all the matrices. 
The 2D and 3D grid sizes are chosen to have comparable matrix sizes. One notes
that the bandwidth reduction is higher in the 2D configuration, which leads
to better performance.  \todo{we do not show performance results in 2D. Do we
need to change anything? Do we have a reference for RCM?} 
This ordering technique is designed to reduce the distance between the 
nonzeros and the diagonal, hopefully allowing a better cache
reutilisation. The reordered version of these two matrices can be seen
in Figure~\ref{fig:rbf2drcm} and in Figure~\ref{fig:rbf3drcm}.

@@

\begin{figure*}
\centering
  \subfigure[Run on the MIC with 240 threads, 64-byte memory alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw}}
%
  \hspace{.3in}
%
  \subfigure[
    Run on the host with 32 threads, 32-byte memory alignment.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}

\caption{Bandwidth performance on the host under idealized conditions as a function of matrix row size.
  Entries with "cpp" denote cases where coding was performed without MIC vector intructions. 
\todo{not 100\% sure what this plots are.}}
\end{figure*}

\begin{figure}
\centering
  \subfigure[Performance of gather and unpack operations on the MIC. ] 
  {\includegraphics[width=.8\linewidth]{figures/bandwidth_gather_unpack.pdf}\label{fig:band_gather}} \\
\end{figure}
%

>>>>

\section{Experimental Validation}
\label{sec:expe}
In this section, we describe a series of numerical experiments to confirm 
the performance predictions of Section~\ref{sec:model}. To better
understand our results, we first perform some bandwidth measurements, followed
by the actual SpMV.

@@

\def\ww{.19\textwidth}
\begin{figure*}[tbh]
  \begin{center}
    \subfigure[Supercompact.]{%
      \label{fig:supercompact}
      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
    \subfigure[Compact.]{%
      \label{fig:compact}
      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
    \subfigure[Random.]{%
      \label{fig:random}
      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
    %% \subfigure[2D, no RCM.]{%
    %%   \label{fig:rbf2dnorcm}
    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
    %% \subfigure[2D, RCM.]{%
    %%   \label{fig:rbf2drcm}
    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
    \subfigure[3D, RCM.]{%
      \label{fig:rbf3dnorcm} 
      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
    \subfigure[3D, RCM.]{%
      \label{fig:rbf3drcm}
      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
  \end{center}

@@

The other two types of matrices are used in derivative stencils in 3D
RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
stencil of a 3D $8^3$ grid). We also apply a Cuthill-Mckee reordering to
all the matrices.  This ordering technique is designed to reduce the
distance between the nonzeros and the diagonal, hopefully allowing a
better cache reutilisation. The reordered version of the matrix can be
seen in Figure~\ref{fig:rbf3drcm}.

@@

In this experimental section, as a point of reference we provide some
results on a typical CPU architecture. This system is ...\todo{include
  a short description of the host.}
All the codes are compiled with icc xxx and -O3 optimization.

\begin{figure*}
\centering
%
 \subfigure[On MIC with 240 threads, 64-byte memory
   alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw_MIC}}
%
  \hspace{.3in}
%
  \subfigure[On CPU with 32 threads, 32-byte memory
    alignment.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:band_rw_CPU}}

  \caption{Bandwidth performance on the host under idealized
    conditions as a function of matrix row size.  Entries with "cpp"
    denote cases where coding was performed without MIC vector
    intructions.  \todo{Can we label the x-axis "number of floats" or
      "size in MB"?}\todo{Can we use the same symbols in both charts?}
    \todo{Can we change the color scheme? in (a), I do not see the
      difference between "read\_cpp" and "read/write cpp"} }
  \label{fig:bandwidth}
\end{figure*}

Figure~\ref{fig:bandwidth} gives some benchmark that compute the
bandwidth achieved when reading and/or writing large arrays on both
the MIC card and the CPU architecture. The read benchmark performs a
simple sum of the array. The write benchmark sets whole the elemtns of
the array to 0. While the read/write benchmarks copies the array into
another one. We also investigate the difference between a straight C++
implementation (compiled with -O3 with proper tagging to inform the
compiler of memory alignments) and a simple implementation using
vector register explicitly.

When using the MIC card (Figure~\ref{fig:band_rw_MIC}) the {\tt
  read\_cpp} implementation reaches a performance of 160GB/s
\todo{gordon check exact number} which is better than the {\tt read}
implementation which only achieves 145GB/s \todo{and that one}. This
is explained by the compiler adding loop unrolling and software
prefetching in the assembly code in the {\tt read\_cpp} case which is
not added in the {\tt read} case. On the other hand, the {\tt write}
case reaches a performance of 155GB/s\todo{and that one} while the
{\tt write\_cpp} implementation never passes 55GB/s. This difference
is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction
which bypasses the Read For Ownership protocol and allows the writes
to be performed in any order. The compiler can not use that
instruction on its own because using would break the memory convention
useful for some parallel algorithms. The {\tt read/write} benchmark
performs at 140GB/s while the {\tt read/write cpp} benchmark performs
at about 90GB/s.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/bandwidth_gather_unpack.pdf}
  \caption{Performance of gather and unpack operations on the MIC.}
  \label{fig:band_gather} 
\end{figure}
%
